---
title: "R Notebook"
output: html_notebook
---

101.5
This is a bit p-hacky, but also prescedented by my Nyvac study. I want to run a logistic regression on every taxon, at every taxonomic level, to find some exemplars of things
that seem to relate to crashes. (adjusting vor multiple comparasons)

I'd like to see if any of those "predictive" (positive sign) bugs increase in abundance as the crash progresses.

```{r}
source("Initial_Processing.R")
library(purrr)
library(broom)
```

```{r}

test_asv <- early %>% filter(ASV == "ASV_4") %>%
  mutate(isCrash = Info1 == "Crash")
test_asv[test_asv$ratio > 0,]$ratio %>% min() -> rat_min
ggplot(data = test_asv, aes(x = log10(ratio), y = Info1)) + geom_point()
test_mod <- glm(isCrash ~ log10(ratio + rat_min), data = test_asv, family = "binomial")
summary(test_mod)
```


```{r}
my_logistic_glm <- function(df){glm(isCrash ~ log10(ratio + rat_min), data = df, family = "binomial")}
quiet_logistic_glm <- quietly(my_logistic_glm)

asv_models <- early %>% 
  mutate(isCrash = Info1 == "Crash") %>%
  group_by(ASV) %>%
  nest() %>%
  mutate(modLs = map(data, quiet_logistic_glm)) %>%
  mutate(mod = map(modLs, ~.[["result"]]),
         warn = map(modLs, ~.[["warnings"]]), # map_chr cant handle zero length vectors
         tid = map(mod, tidy)
         ) %>%
  identity()
```

```{r}
asv_model_data <- asv_models %>%
  select(ASV, tid) %>%
  unnest(tid) %>%
  filter(term != "(Intercept)")

asv_model_ok <- asv_model_data %>%
  filter(`p.value` < 0.05)
asv_model_ok
```
Hah! Nothing comes close to predicting crashes.

# What about after the crashes

```{r}
many_logistic <- function(fort, gb = ASV){
  asv_models <- fort %>% 
  mutate(isCrash = Info1 == "Crash") %>%
  group_by({{gb}}) %>% # or gtax
  nest() %>%
  mutate(modLs = map(data, quiet_logistic_glm)) %>%
  mutate(mod = map(modLs, ~.[["result"]]),
         warn = map(modLs, ~.[["warnings"]]), # map_chr cant handle zero length vectors
         tid = map(mod, tidy)
         ) %>%
  identity()
  
  asv_model_data <- asv_models %>%
  select({{gb}}, tid) %>%
  unnest(tid) %>%
  filter(term != "(Intercept)")
}
```

```{r}
asv_model_group_test <- many_logistic(early, Family)
asv_model_group_test %>% arrange(p.value) %>% filter(p.value < 0.05)
```
Now, somehow there are 11 statistically significant families.
I don't get it. When I group to all levels, I don't get this result, but when I just group to family, it looks like this. What's going on?
One thing is that I'm not doing a summarize call first.

```{r}
asv_model_data_01 <- many_logistic(early)
asv_model_data_01 %>% filter(p.value < 0.05)
```

```{r}
asv_model_data_mid <- many_logistic(mid)
asv_model_data_mid_ok <- asv_model_data_mid %>% filter(p.value < 0.05)
asv_model_data_mid_ok
```

One weak thing.

# What about by different taxonomic levels

Test group taxa

```{r}
group_taxa_2(early, "Family")
```


```{r}
taxLevels <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")
glomThing <- tibble(taxLevel = taxLevels) %>%
  mutate(data1 = map(taxLevel, ~group_taxa_2(early, .)))
```

```{r}
glomThing_2 <- glomThing %>% 
  mutate(data2 = map(data1, many_logistic)) %>% # I should modify many_logistic to group by the named level
  unnest(data2) %>%
  select(-data1)
```

Ok this is silly, everything is named by the first asv in the group still, but should still work-ish

Something isn't making sense in the output here. Why aren't there coherent groups?

```{r}
glomThing_2 %>% filter(p.value < 0.05)
```
Nothing

# And mid?

```{r}
taxLevels <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")
glomThing <- tibble(taxLevel = taxLevels) %>%
  mutate(data1 = map(taxLevel, ~group_taxa_2(mid, .)))
```
# Debug here and in many logistic
```{r}
glomThing_2 <- glomThing %>% 
  mutate(data2 = map2(data1, taxLevel, many_logistic)) %>%
  unnest(data2) %>%
  select(-data1)
```

Ok this is silly, everything is named by the first asv in the group still, but should still work-ish

```{r}
glomThing_2 %>% filter(p.value < 0.05)
```

Ok. Can't even p-hack my way out of this non-finding.