---
title: "R Notebook"
output: html_notebook
---

So I think the first time-point isn't actually correlated with survival after all. I may have been mistreding my statistics output.
In any case I'd like to show how the microbiota do vary over the life of a crash.

```{r}
library(ggvegan)
library(plotly)
library(purrr)
library(broom)
library(flextable)
source("Initial_Processing.R")
```

## Canonical correspondence analysis.
Overall patterns in community structure over the life time of things.

```{r}
#first_cca <- cca(sqrt(ratioDf) ~ DayN + Info1, data = metadata_main )
first_cca <- cca(sqrt(ratioDf) ~ DayN + Info1, data = metadata_main )
set.seed(33100)
anova(first_cca, by = "margin")
plot(first_cca)
```
Ok. Lets look at how the samples evolve over time on the above plot



```{r}
first_cca_wa <- fortify(first_cca, display = "wa") %>% 
  #rename(Label = label) %>% # during an update of ggvegan "Label" was changed to "label"
  left_join(metadata_main, by = c("label" = "Sample0")) %>%
  mutate(DayCat = ordered(DayCat, levels = unique(DayCat)))


early_days <- first_cca_wa %>% filter(DayN < 14) %>% mutate(DayCat = droplevels(DayCat))
late_days <- first_cca_wa %>% filter(DayN >= 14)
plt1 <- early_days %>% 
  ggplot(aes(x = CCA1, y = CCA2, shape = DayCat, group = `Brood#`, color = Info1)) +
  geom_path(aes(x = CCA1, y = CCA2, color = Info1, group = `Brood#`), data = first_cca_wa, inherit.aes = FALSE) +
  geom_point(size = 3, stroke = 1.5) + 
  scale_shape_manual(values = c(21, 22, 23, 1, 1, 1, 1)) +
  #scale_shape_manual(values = c(21,22,23), breaks = c("3-5", "7-8", "10-12")) +
  #scale_fill_manual(values = rep(c("green", "blue", "black"), 5)) +
  #scale_color_manual(values = c(Good = "black", Crash = "red")) +
  scale_color_manual(values = c(Good = "darkblue", Crash = "red")) +
  geom_text(data = late_days, aes(x = CCA1, y = CCA2, label = DayCat), inherit.aes = FALSE) +
  geom_text(data = first_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good") %>% mutate(label = str_remove(label, "N")), inherit.aes = FALSE, aes(x = CCA1 * 3.4, y = CCA2 * 3.4, label = label), size = 5) +
  geom_segment(data = first_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good"), inherit.aes = FALSE, aes(x = 0, y = 0, xend = CCA1 * 3.2, yend = CCA2 * 3.2), color = "brown4",
               arrow = arrow(length = unit(0.3, "cm")), size = 1) +
  geom_label(data = first_cca %>% fortify(display = "cn") %>% mutate(label = str_remove(label, "Info1")), inherit.aes = FALSE, aes(x = CCA1, y = CCA2, label = label), alpha = 0.5, size = 5) +
  theme_bw(base_size = 16)
plt1
```


# As above, but dropping the days after day 12

```{r}
metadata_limited <- metadata_main %>% filter(DayN < 14)
ratioDf_limited <- ratioDf[rownames(ratioDf) %in% metadata_limited$Sample0,]
second_cca <- cca(sqrt(ratioDf_limited) ~ DayN + Info1, data = metadata_limited)
anova(second_cca, by = "margin")
```
```{r}
Info2_Levels <- c("Best", "Ok", "Poor", "NoSurvivors")
```



```{r}
larvae_cca_plotter_limited <- function(second_cca){
second_cca_wa <- fortify(second_cca, display = "wa") %>% 
  left_join(metadata_main, by = c("label" = "Sample0")) %>% 
  mutate(DayCat = ordered(DayCat, levels = unique(DayCat))) %>%
  identity()


early_days <- second_cca_wa %>% filter(DayN < 14) %>% mutate(DayCat = droplevels(DayCat))
late_days <- second_cca_wa %>% filter(DayN >= 14)
plt2 <- early_days %>% 
  ggplot(aes(x = CCA1, y = CCA2, shape = DayCat, group = `Brood#`, color = Info1)) +
  geom_path(aes(x = CCA1, y = CCA2, color = Info1, group = `Brood#`), data = second_cca_wa, inherit.aes = FALSE) +
  geom_point(size = 3, stroke = 1.5) + 
  scale_shape_manual(values = c(1, 2, 7, 1, 1, 1, 1)) +
  #scale_shape_manual(values = c(21,22,23), breaks = c("3-5", "7-8", "10-12")) +
  #scale_fill_manual(values = rep(c("green", "blue", "black"), 5)) +
  #scale_color_manual(values = c(Good = "black", Crash = "red")) +
  scale_color_manual(values = c(Good = "darkblue", Crash = "red")) +
  geom_text(data = late_days, aes(x = CCA1, y = CCA2, label = DayCat), inherit.aes = FALSE) +
  geom_text(data = second_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good") %>% mutate(label = str_remove(label, "N")), inherit.aes = FALSE, aes(x = CCA1 * 2.4, y = CCA2 * 2.4, label = label), size = 5) +
  geom_segment(data = second_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good"), inherit.aes = FALSE, aes(x = 0, y = 0, xend = CCA1 * 2.2, yend = CCA2 * 2.2), color = "blue",
               arrow = arrow(length = unit(0.3, "cm")), size = 1) +
  geom_label(data = second_cca %>% fortify(display = "cn") %>% mutate(label = str_remove(label, "Info1")), inherit.aes = FALSE, aes(x = CCA1, y = CCA2, label = label), alpha = 0.5, size = 4) +
  theme_bw(base_size = 16)
plt2
}

larvae_cca_plotter_limited(second_cca)
```


```{r}
ggsave(here("Figures", "CCA_Days3-12.png"), plot = plt2, height = 4, width = 6)
```



# Check each day

```{r}
dayCatLev <- c("3-5", "7-8", "10-12")
smaller_ccas <- metadata_main %>%
  filter(DayCat %in% dayCatLev) %>%
  group_by(DayCat) %>%
  nest() %>%
  mutate(rdf = map(data, ~ratioDf[rownames(ratioDf) %in% .$Sample0,])) %>%
  mutate(mod = map2(.x = data, .y = rdf, ~cca(sqrt(.y) ~  Info1, data = .x))) %>%
  mutate(anova = map(mod, anova)) %>%
  mutate(anova_table = map(anova, tidy))

smaller_ccas2 <- smaller_ccas %>%
  select(DayCat, anova_table) %>%
  unnest(anova_table)
smaller_ccas2

smaller_ccas2 %>% filter(term == "Model") %>% left_join(
smaller_ccas2 %>% pivot_wider(id_cols = DayCat, names_from = term, values_from = ChiSquare) %>%
  mutate(PercentChi = Model/(Model + Residual)),
by = "DayCat"
) %>%
  select(-c(Model, term, df))
```

In short though, we are no effect on the pre-crash, but the community does change during the crash.


# What about at different taxonomic levels?

Problem with quasiquotation. I need to specify the level with a string, but I cant `enquo` a string and I'm kind of stuck.

```{r}
taxLevels <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Tag_ASV")

# group_by_taxlevel <- function(taxlevel, data = nonOyster_wm %>% filter(Info1 %in% c("Crash", "NoCrash"))){
#   taxlevelQ <- enquo(taxlevel)
#   data %>%
#     group_by(Sample, `Brood#`, DayN, Type, Info1, Info2, !!taxlevelQ) %>%
#     summarise(ratio = sum(ratio)) %>%
#     ungroup()
#}

# was nonOyster_wm # then nonOyster20_wm
group_by_taxlevel <- function(taxlevel, data = nonOyster_wm %>% filter(Info1 %in% c("Crash", "Good"))){
  taxlevelQ <- enquo(taxlevel)
  data %>%
    group_by(Sample, `Brood#`, DayN, Type, Info1, Info2, "Taxid" = rlang::eval_tidy(sym(taxlevel))) %>%
    summarise(ratio = sum(ratio), n = n()) %>%
    ungroup()
}

testGrp <- group_by_taxlevel("Tag_ASV")
```


```{r}
metabeast <- tibble(taxlevel = taxLevels) %>%
  # group by taxonomic level
  mutate(df = map(taxlevel, group_by_taxlevel)) %>%
  # pivot each grouping wider
  mutate(comm = map(df, function(x){x %>% select(Sample, Taxid, ratio) %>% pivot_wider(names_from = Taxid, values_from = ratio)})) %>%
  mutate(comm = map(comm, ~column_to_rownames(., "Sample"))) %>%
  identity()



metabeast2 <- expand_grid(taxlevel = taxLevels, daycat = dayCatLev) %>% left_join(metabeast) %>%
  mutate(meta = map(daycat,  ~filter(metadata_main, DayCat == .))) %>%
  #mutate(comm2 = map2(.x = comm, .y = meta, ~.x[rownames(.x) %in% .y$Sample0,]))
  identity()

comm_subsettor <- function(comm, meta){
   # print(comm)
   # print(meta)
  comm[rownames(comm) %in% meta$Sample0,]
}

#comm_subsettor(metabeast2$comm[[4]], metabeast2$meta[[4]])


metabeast3 <- metabeast2 %>%
  mutate(comm2 = map2(.x = comm, .y = meta, ~comm_subsettor(comm = .x, meta = .y)))
```

```{r}
# make sure that a sample shows up at least n times in the sample of interest
# lets say 20% of samples
shows_up_enough <- function(cdf, nn = NULL, pct = NULL){
  howoften <- apply(cdf > 0, 2, sum)
  # cdf1 <- case_when(
  #   (is.null(nn) & !is.null(pct)) ~  cdf[,howoften >= (dim(cdf)[1] * pct)],
  #   (!is.null(nn) & is.null(pct)) ~  cdf[,howoften >= nn]#,
  #   #TRUE ~ stop("either but not both of n and pct must be defined") # why is this reaching true?
  # )
  if(is.null(nn) & !is.null(pct)){
    cdf1 <- cdf[,howoften >= (dim(cdf)[1] * pct)]
  } else if(!is.null(nn) & is.null(pct)) {
    cdf1 <- cdf[,howoften >= nn]
  }else{
    stop("either but not both of n and pct must be defined")
  }
  
  return(cdf1)
}

metabeast3b <- metabeast3 %>%
  mutate(comm3 = map(comm2, ~shows_up_enough(.,n = 3)))
```


```{r}

# on to cca
metabeast4 <- metabeast3b %>%
  mutate(cca = map2(.x = comm3, .y = meta, ~cca(sqrt(.x) ~ Info1, data = .y)))

#Yes!

metabeast5 <- metabeast4 %>%
  mutate(anova = map(cca, ~anova(., permutations = how(nperm = 9999)))) %>%
  mutate(tanova = map(anova, tidy))

metabeast_table <- metabeast5 %>%
  select(taxlevel, daycat, tanova) %>%
  unnest(tanova) %>%
  filter(term == "Model")
  # flextable() %>%
  # bold()
```
So the strange things is that the TAG_ASV isn't matching the run I did earlier.
However, I do 

To do: anova all of these and make into one table.

```{r}
metabeast_flex <- metabeast_table %>%
  mutate(daycat = ordered(daycat, levels = dayCatLev),
         taxlevel = ordered(taxlevel, levels = taxLevels)) %>%
  mutate(FDR = p.adjust(`p.value`)) %>%
  mutate(taxlevel = recode(taxlevel, "Tag_ASV" = "ASV")) %>%
  select(daycat, everything()) %>%
  arrange(daycat, taxlevel) %>%
  #mutate(ChiSquare = prettyNum(ChiSquare, digits = 2)) %>%
  mutate(across(ChiSquare:FDR, ~formatC(.x, digits = 2, format = "g", flag = "#"))) %>%
  rename(`Age (Days)` = daycat, DF = df, `Chi Square` = ChiSquare, F = "statistic") %>%
  rename(p = p.value) %>%
  #mutate(across(statistic, ~prettyNum(.x))) %>%
  flextable() %>%
  bold(i = ~ `p` < 0.05, j = "p") %>% merge_v(j = 1) %>%
  bold(i = ~ FDR < 0.2, j = "FDR") %>%
  theme_vanilla() %>%
  identity()
metabeast_flex
```

I'm confused about why the 10-12 tag asv samples don't look like they do earlier in this notebook.
[These should differ with the different grouping, but aren't. Check this.]

```{r}
flextable::save_as_docx(metabeast_flex, path = here("Tables", "Metabeast.docx"))
flextable::save_as_html(metabeast_flex, path = here("Tables", "Metabeast.html"))
flextable::save_as_image(metabeast_flex, path = here("Tables", "Metabeast.png"))
flextable::save_as_pptx(metabeast_flex, path = here("Tables", "Metabeast.pptx"))
```



Checking that thing are the same
```{r}
smaller_ccas %>% select(DayCat, rdf) %>%
  mutate(rows = map_int(rdf, nrow), cols = map_int(rdf, ncol)) %>%
  select(DayCat, rows, cols)
```

```{r}
metabeast5 %>% select(taxlevel, daycat, comm2)
```
These are very different numbers of rows
I think I was subsetting to present in 20 before
Did that and now nothing is statistically significant (which does't match the earlier work).
I'm so confused about what is going on.

```{r}
test_mb <- metabeast5[(metabeast5$taxlevel == "Tag_ASV" & metabeast5$daycat == "3-5"),]$comm2[[1]]
test_sm <- smaller_ccas[smaller_ccas$DayCat == "3-5",]$rdf[[1]]
test_mb_meta <- metabeast5[(metabeast5$taxlevel == "Tag_ASV" & metabeast5$daycat == "3-5"),]$meta[[1]]
```

these appear to be identical data frames with columns in different orders. I don't understand why the results would differ.
My hunch is that something is going on strange over in metabeast land, and hense the nothing is significant thing.

The problem may be that there is no clear join operator on the metadata. How does it know which row is which, especially if the order differs at all.

Nope, order is identical
> rownames(test_mb)
 [1] "H144" "H158" "H207" "H209" "H219" "H221" "H222" "H224" "H229" "H237" "H259" "H303" "H40"  "H48"  "H94" 
> test_mb_meta$Sample0
 [1] "H144" "H158" "H207" "H209" "H219" "H221" "H222" "H224" "H229" "H237" "H259" "H303" "H40"  "H48"  "H94" 
 
 > smaller_ccas[3,]$mod
[[1]]
Call: cca(formula = sqrt(.y) ~ Info1, data = .x)

              Inertia Proportion Rank
Total          0.8716     1.0000     
Constrained    0.1260     0.1446    1
Unconstrained  0.7456     0.8554   13
Inertia is scaled Chi-square 

Eigenvalues for constrained axes:
   CCA1 
0.12602 

Eigenvalues for unconstrained axes:
    CA1     CA2     CA3     CA4     CA5     CA6     CA7     CA8     CA9    CA10    CA11    CA12    CA13 
0.16890 0.15594 0.10945 0.06156 0.05399 0.04816 0.03699 0.02898 0.02550 0.01819 0.01493 0.01163 0.01138 


> metabeast5[(metabeast5$taxlevel == "Tag_ASV" & metabeast5$daycat == "10-12"),]$cca[[1]]
Call: cca(formula = .x ~ Info1, data = .y)

              Inertia Proportion Rank
Total         1.36420    1.00000     
Constrained   0.10720    0.07858    1
Unconstrained 1.25700    0.92142   13
Inertia is scaled Chi-square 

Eigenvalues for constrained axes:
  CCA1 
0.1072 

I forgot to squrt transform the metabeast

Strangely, everything is statistically significant if I include rare ASVs.
I need to look up what rare bugs do to cca.

## Example to share

```{r}
save(test_mb, test_mb_meta, file = "test_mb.RData")
```


```{r}
cca(sqrt(test_mb) ~ Info1, data = test_mb_meta) %>% anova(nperm = 9999)
```
```{r}
howoften <- apply(test_mb > 0, 2, sum)
test_mb_20<- test_mb[,howoften >= 2]
```

```{r}
cca(sqrt(test_mb_20) ~ Info1, data = test_mb_meta) %>% anova(nperm = 9999)
```

????
Well, that blows a hole in that hypothesis. 
This comes up as significant even when I do subset it.
Why doesn't any of my code make sense?
Somehow test_mb_20 is still way bigger than test_sm
and I don't understand why.

Its because I was only taking ASVs that showed up at least three times once.

Before I was at least 20% of samples

```{r}
howoften <- apply(test_mb > 0, 2, sum)
test_mb_20<- test_mb[,howoften >= (dim(test_mb)[1] * 0.2)]
```

the challenge is that the test is in presence abscence and I've removed count data.
What if, to make my life easier, I just do things that have shown up in at least 3 samples.

So, I get a significant result if I include singletons, but not if I remove them.

# Redo CCA with family level patterns.
cca(sqrt(ratioDf) ~ DayN + Info1, data = metadata_main )
metabeast4 <- metabeast3b %>%
  mutate(cca = map2(.x = comm3, .y = meta, ~cca(sqrt(.x) ~ Info1, data = .y)))
```{r}
metawhole <- metabeast %>%
  mutate(cca = map(.x = comm, ~cca(sqrt(.x) ~ Info1 + DayN, data = metadata_main)))
```

```{r}
family_cca <- metawhole %>% filter(taxlevel == "Family") %>% pull(cca) %>% .[[1]]
```

```{r}
family_cca_wa <- fortify(family_cca, display = "wa") %>% left_join(metadata_main, by = c("label" = "Sample0")) %>%
  mutate(DayCat = ordered(DayCat, levels = unique(DayCat)))
```

```{r}
early_days_fam <- family_cca_wa %>% filter(DayN < 14) %>% mutate(DayCat = droplevels(DayCat))
late_days_fam <- family_cca_wa %>% filter(DayN >= 14)
plt3 <- early_days_fam %>% 
  ggplot(aes(x = CCA1, y = CCA2, shape = DayCat, group = `Brood#`, color = Info2)) +
  geom_path(aes(x = CCA1, y = CCA2, color = Info2, group = `Brood#`), data = family_cca_wa, inherit.aes = FALSE) +
  geom_point(size = 3, stroke = 1.5) + 
  scale_shape_manual(values = c(21, 22, 23, 1, 1, 1, 1)) +
  #scale_shape_manual(values = c(21,22,23), breaks = c("3-5", "7-8", "10-12")) +
  scale_fill_manual(values = rep(c("green", "blue", "black"), 5)) +
  #scale_color_manual(values = c(Good = "black", Crash = "red")) +
  scale_color_manual(values = c(Best = "black", Ok = "blue", Poor = "Orange", NoSurvivors = "red")) +
  geom_text(data = late_days_fam, aes(x = CCA1, y = CCA2, label = DayCat), inherit.aes = FALSE) +
  geom_text(data = family_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good") %>% mutate(label = str_remove(label, "N")), inherit.aes = FALSE, aes(x = CCA1 * 3.4, y = CCA2 * 3.4, label = label), size = 5) +
  geom_segment(data = family_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good"), inherit.aes = FALSE, aes(x = 0, y = 0, xend = CCA1 * 3.2, yend = CCA2 * 3.2), color = "brown4",
               arrow = arrow(length = unit(0.3, "cm")), size = 1) +
  geom_label(data = family_cca %>% fortify(display = "cn") %>% mutate(label = str_remove(label, "Info1")), inherit.aes = FALSE, aes(x = CCA1, y = CCA2, label = label), alpha = 0.5, size = 5) +
  theme_bw(base_size = 16)
plt3
```

```{r}
metadata_limited <- metadata_main %>% filter(DayN < 14)
ratioDf_limited <- ratioDf[rownames(ratioDf) %in% metadata_limited$Sample0,]
second_cca <- cca(sqrt(ratioDf_limited) ~ DayN + Info1, data = metadata_limited)
anova(second_cca, by = "margin")
```

```{r}
minrat <- min(ratioDf_limited[ratioDf_limited>0])
second_rda <- rda(log(ratioDf_limited + minrat) ~ DayN + Info1, data = metadata_limited)
anova(second_rda, by = "margin")
```


```{r}
metawhole <- metabeast %>%
  mutate(cca = map(.x = comm, ~cca(sqrt(.x) ~ Info1 + DayN, data = metadata_main)))
```

ratioDf_limited <- ratioDf[rownames(ratioDf) %in% metadata_limited$Sample0,]
```{r}
metawhole2 <- metabeast %>%
  mutate(comm_limited = map(comm, ~.[rownames(.) %in% metadata_limited$Sample0,])) %>%
  mutate(cca_limited = map(comm_limited, ~cca(sqrt(.) ~ Info1 + DayN, data = metadata_limited)))
```

```{r}
family_cca_limited <- metawhole2[metawhole2$taxlevel == "Family","cca_limited"][[1]][[1]]
```


```{r}
larvae_cca_plotter_limited <- function(second_cca){
second_cca_wa <- fortify(second_cca, display = "wa") %>% 
  left_join(metadata_main, by = c("label" = "Sample0")) %>% 
  mutate(DayCat = ordered(DayCat, levels = unique(DayCat))) %>%
  identity()
}

larvae_cca_plotter_limited(family_cca_limited)
```
family

## For proposal
```{r}
larvae_cca_plotter_limited_prop <- function(second_cca){
second_cca_wa <- fortify(second_cca, display = "wa") %>% 
  left_join(metadata_main, by = c("label" = "Sample0")) %>% 
  mutate(DayCat = ordered(DayCat, levels = unique(DayCat))) %>%
  identity()


early_days <- second_cca_wa %>% filter(DayN < 14) %>% mutate(DayCat = droplevels(DayCat))
late_days <- second_cca_wa %>% filter(DayN >= 14)
plt2 <- early_days %>% 
  ggplot(aes(x = CCA1, y = CCA2, shape = DayCat, group = `Brood#`, color = Info1)) +
  geom_path(aes(x = CCA1, y = CCA2, color = Info1, group = `Brood#`), alpha = 0.5, data = second_cca_wa, inherit.aes = FALSE) +
  geom_point(size = 3, stroke = 1.5, alpha = 0.7) + 
  scale_shape_manual(values = c(1, 2, 7, 1, 1, 1, 1), name = "Larval Age (Days)") +
  #scale_shape_manual(values = c(21,22,23), breaks = c("3-5", "7-8", "10-12")) +
  #scale_fill_manual(values = rep(c("green", "blue", "black"), 5)) +
  scale_color_manual(values = c(Good = "darkblue", Crash = "red"), name = "Batch Outcome") +
  #scale_color_manual(values = c(Best = "black", Ok = "blue", Poor = "Orange", NoSurvivors = "red")) +
  geom_text(data = late_days, aes(x = CCA1, y = CCA2, label = DayCat), inherit.aes = FALSE) +
  geom_text(data = second_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good") %>% mutate(label = str_remove(label, "N")) %>% mutate(label = str_replace(label, "Day", "Age")), inherit.aes = FALSE, aes(x = CCA1 * 2.4, y = CCA2 * 2.4, label = label), size = 5) +
  geom_segment(data = second_cca %>% fortify(display = "bp") %>% filter(label != "Info1Good"), inherit.aes = FALSE, aes(x = 0, y = 0, xend = CCA1 * 2.2, yend = CCA2 * 2.2), color = "darkgreen",
               arrow = arrow(length = unit(0.3, "cm")), size = 1) +
  geom_label(data = second_cca %>% fortify(display = "cn") %>% mutate(label = str_remove(label, "Info1")), inherit.aes = FALSE, aes(x = CCA1, y = CCA2, label = label), alpha = 0.5, size = 4) +
  theme_bw(base_size = 16) +
  theme(legend.position = c(.225, .65), legend.background = element_blank())

plt2
}

pre_prop_fig1 <- larvae_cca_plotter_limited_prop(family_cca_limited)
pre_prop_fig1
```
```{r}
ggsave(here("Figures", "Pre_Prop_Fig1.png"), pre_prop_fig1, height = 5, width = 6)
```

I wonder if I could add families scores to this latest plot.

```{r}
family_cca %>% fortify(display = "species")
```

But really I just want frequent families.

